A decision support tool for the Home Health Care Routing and Scheduling Problem

### 2022-01-25 tex/Paper/main.tex
added pseudocode.

%----------------------------------------------------------------------------------%
\subsection{Shift Work}
\label{sub:shiftwork}
\noindent There may be cases where some nurses have multiple shifts in a single day, e.g. a morning shift and an evening shift; these nurses will not be available at all times during the day. Let $\Pi_i = \{\pi_1^i,\dotsc, \pi_{g_i}^i\}$ be the set of shifts for a nurse $i \in N$, where $g_i$ is the total number of shifts for nurse $i$ during the day. For each shift $\pi_f^i \in \Pi_i$, we denote the shift start time by $u(\pi_f^i)$ and the shift end time by $v(\pi_f^i)$. Therefore, the total contracted working time, $ct(i)$, for a nurse $i$ is calculated as:
\begin{equation}
	ct(i) = \displaystyle\sum_{f=1}^{g_i} v(\pi_f^i) - u(\pi_f^i). \label{eq:contractedworkingtime}
\end{equation}
\noindent Clearly, if $g_i = 1$ for a nurse $i \in N$, then the nurse only has one shift; thus $ct(i) = en(i)-st(i)$. Given these shifts, we must ensure that jobs are assigned to nurses during their actual working time. This also involves considering two cases that may occur:
\begin{enumerate}[label={(\arabic*)}]
	%\item A nurse $i$ is assigned a job $j$ with time window within the nurse's valid shift time, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period. In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the beginning of the nurse's next shift, and the start time for $j$ is moved forward to the start of the next shift. The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
	%\item A nurse $i$ is assigned a job $j$ with time window within nurse $i$'s break period. In this case, the waiting time incurred is the entire break period, and the arrival time at $j$ is moved forward to the start of the next shift. Tardiness as calculated as above.
	\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period. In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the end of the shift, and the start time for $j$ is moved forward to the start of $i$'s next shift. The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time. In this case, no waiting time is incurred and the new start time for job $j$ and tardiness are as above.
\end{enumerate}
\begin{figure}[h!]
	\centering	
	\begin{subfigure}[h]{\textwidth}
		\centering
		\includestandalone[width=0.7\textwidth]{figures/shiftwork1}
		\vspace{-2mm}\caption{}
		\label{fig:itemsinfeas}
		\vspace{2mm}
	\end{subfigure}
	\begin{subfigure}[h]{\textwidth}
		\centering
		\includestandalone[width=0.7\textwidth]{figures/shiftwork2}
		\vspace{-2mm}\caption{}
		\label{fig:itemsfeas}
	\end{subfigure}
	\caption{Examples of the alteration of a job $j$ depending on a nurse $i$'s shift times according to the scenarios above, when (a) the duration of $j$ extends past the end of $i$'s shift; and (b) the start time of $j$ occurs during a period of unavailable time.}
	\label{fig:itemsaligned}
\end{figure}
%\correction{What happens if the next "available" slot is not enough for the job.}
\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule. For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid. Then, the waiting times will be corrected by removing all relevant breaks, as shown in \eqref{eq:waitingtimenurse}.

>>>>

%----------------------------------------------------------------------------------%
\subsection{Shift Work}
\label{sub:shiftwork}
\noindent There may be cases where some nurses have multiple shifts in a single day, e.g. a morning shift and an evening shift; these nurses will not be available at all times during the day. Let $\Pi_i = \{\pi_1^i,\dotsc, \pi_{g_i}^i\}$ be the set of shifts for a nurse $i \in N$, where $g_i$ is the total number of shifts for nurse $i$ during the day. For each shift $\pi_f^i \in \Pi_i$, we denote the shift start time by $u(\pi_f^i)$ and the shift end time by $v(\pi_f^i)$. Therefore, the total contracted working time, $ct(i)$, for a nurse $i$ is calculated as:
\begin{equation}
	ct(i) = \displaystyle\sum_{f=1}^{g_i} v(\pi_f^i) - u(\pi_f^i). \label{eq:contractedworkingtime}
\end{equation}
\noindent Clearly, if $g_i = 1$ for a nurse $i \in N$, then the nurse only has one shift; thus $ct(i) = en(i)-st(i)$. Given these shifts, we must ensure that jobs are assigned to nurses during their actual working time. This also involves considering two cases that may occur:
\begin{enumerate}[label={(\arabic*)}]
	%\item A nurse $i$ is assigned a job $j$ with time window within the nurse's valid shift time, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period. In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the beginning of the nurse's next shift, and the start time for $j$ is moved forward to the start of the next shift. The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
	%\item A nurse $i$ is assigned a job $j$ with time window within nurse $i$'s break period. In this case, the waiting time incurred is the entire break period, and the arrival time at $j$ is moved forward to the start of the next shift. Tardiness as calculated as above.
	\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period. In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the end of the shift, and the start time for $j$ is moved forward to the start of $i$'s next shift. The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time. In this case, no waiting time is incurred and the new start time for job $j$ and tardiness are as above.
\end{enumerate}
\begin{figure}[h!]
	\centering	
	\begin{subfigure}[h]{\textwidth}
		\centering
		\includestandalone[width=0.7\textwidth]{figures/shiftwork1}
		\vspace{-2mm}\caption{}
		\label{fig:itemsinfeas}
		\vspace{2mm}
	\end{subfigure}
	\begin{subfigure}[h]{\textwidth}
		\centering
		\includestandalone[width=0.7\textwidth]{figures/shiftwork2}
		\vspace{-2mm}\caption{}
		\label{fig:itemsfeas}
	\end{subfigure}
	\caption{Examples of the alteration of a job $j$ depending on a nurse $i$'s shift times according to the scenarios above, when (a) the duration of $j$ extends past the end of $i$'s shift; and (b) the start time of $j$ occurs during a period of unavailable time.}
	\label{fig:itemsaligned}
\end{figure}
%\correction{What happens if the next "available" slot is not enough for the job.}
%\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
\noindent The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$. For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job. Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e. $g_i = 1$. Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ and job $j$ would be within a break. In both of these cases, the function \textsc{FindValidTime} is used. This function scans through the nurse's day to find the next available time period in which the nurse can feasibly perform the job $j$, and returns the start time of this period, $t_j$. If the job $j$ is a DS, then \textsc{FindValidTime} seeks a start time that is suitable for the current nurse $i$ and the second nurse on the job.

\begin{algorithm}[h]
\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
\label{alg:shiftwork}
\begin{algorithmic}[1]
	\State Let $f$ be the current shift for nurse $i$
	\State $a(i,0) \gets 0$
	\For{$j \gets 1 \To m_i$}
		\State $a(i,0) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
		\If{$g_i = 1$}
			\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
			\State $z_j \gets \max\{0, (a(i,j) - l_j))\}$
			\State $t_j \gets a(i,j) + w_{ij}$
		\ElsIf{$g_i \neq 1 \AAnd \max\{e_j + s_j, a(i,j) + s_j\} > v(\pi_{f}^{i})$}
			\State $t_j \gets$ \textsc{FindValidTime}$(i, j)$
			\State $w_{ij} \gets t_j - l_j$
			\State $z_j \gets t_j - l_j$
			\State $f \gets$ shift containing time $t_j$
		\ElsIf{$g_i \neq 1 \AAnd a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f}^{i})$}
			\State $t_j \gets$ \textsc{FindValidTime}$(i,j)$
			\State $w_{ij} \gets t_j - v(\pi_{f}^{i})$
			\State $z_j \gets t_j - l_j$
			\State $f \gets$ shift containing time $t_j$
		\EndIf
	\EndFor
\end{algorithmic}
\end{algorithm}

For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid. Then, the waiting times will be corrected by removing all relevant breaks, as shown in \eqref{eq:waitingtimenurse}.
